Exploring Design Space for Differentially Private Bayesian Inference

### 2019-01-03 ICML2019/main.tex
pass on section 2

\begin{abstract} Bayesian inference is a statistical method which
allows one to derive a \emph{posterior} distribution,
starting from a \emph{prior} distribution and observed data. Several
approaches have been explored in order to make this process differentially
private. For example, \cite{dimitrakakis2014robust}, and
\cite{wang2015privacy} proved that, under specific conditions, sampling from
the posterior distribution is already differentially private. \cite{zhang2016differential}, \cite{foulds2016theory},
designed differentially private mechanisms that output a
representation of the full posterior distribution.

@@

Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression $\Pr(\dataobs | \theta)$ denotes the
\emph{likelihood} of observing $\dataobs$ under a value of
$\theta$. Since we consider $\dataobs$ to be fixed, the likelihood is
a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$, that is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate to the likelihood function. Use of a conjugate prior
simplifies calculations and allows for inference to be performed in a
recursive fashion over the data. Then, we have:
\[
  \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]

@@

In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
specifically, a Beta-binomial mode. We will consider the situation the underlying data is binomial distribution ($\thicksim binomial(\theta)$), where $\theta$ represents
the parameter --informally called \emph{bias}-- of a Bernoulli
distributed random variable. The
prior distribution over $\theta\in [0,1]$ is going to be a beta
distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this, the inference can be easily derived:
\[
   \bysinfer(binomial(\theta), \betad(\alpha, \beta), \dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]

@@

The beta-binomial model can be immediately generalized to Dirichlet-multinomial, with underlying data multinomially distributed. The \emph{bias} is represented by parameter $\vtheta$, the vector of parameters of a categorically distributed random variable. The prior distribution over $\vtheta\in [0,1]^{k}$
is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbb{R}^{+})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where $\iverson{\cdot}$ represents Iverson bracket notation.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots \Delta\alpha_k)$ the inference over $\vtheta$ turns out to be
\[
   \bysinfer(multinomial(\theta), \dirichlet(\valpha), \dataobs)=\dirichlet(\valpha + \Delta \valpha)
\]
where $+$ denotes the componentwise sum of vectors of reals.

@@

A randomized mechanism $\mathcal{M}$ with domain $\mathbb{N}^{|\mathcal{X}|}$ is $(\epsilon, \delta)-$differential privacy, if for all adjacent
\footnote{Given $\dataobs, \dataobs'$  we say that $\dataobs$ and $\dataobs'$ are adjacent and we write, $\adj{\dataobs}{\dataobs'}$, iff\\
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$. } $\dataobs, \dataobs' \in \mathbb{N}^{|\mathcal{X}|}$ and all $\mathcal{S} \subset Range(\mathcal{M})$:
\begin{equation*}
\Pr[\mathcal{M}(\dataobs) \in \mathcal{S}] = e^\epsilon \Pr[\mathcal{M}(\dataobs') \in \mathcal{S}].
\end{equation*}

@@

However, in previous works, the accuracy of the posterior was measured again with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as defined above. This accuracy metric is meaningless when the results released are distributions rather than numerical values. In contrast, distribution metrics such as $f$-divergence, Hellinger distance, etc. come into mind overtly when we are measuring distance between distributions. This gives us motivation on exploring the design space of mechanisms by considering different norms (a distribution metric) to compute the sensitivity
and provide guarantees on the accuracy.

Specifically, we will use the Hellinger distance $\hellinger(\cdot,\cdot)$:
Given two beta distributions
$\betad(\alpha_1, \beta_1),$ and $\betad(\alpha_2, \beta_2)$ the following equality holds 
\[
  \hellinger(\betad(\alpha_1, \beta_1), \betad(\alpha_2, \beta_2))=
  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]

>>>>

\begin{abstract} Bayesian inference is a statistical method which
allows to derive a \emph{posterior} distribution,
starting from a \emph{prior} distribution and observed data. Several
approaches have been explored in order to make this process differentially
private. For example \cite{dimitrakakis2014robust} and
\cite{wang2015privacy} proved that, under specific conditions, sampling from
the posterior distribution is already differentially private, while \cite{zhang2016differential} and \cite{foulds2016theory}
designed differentially private mechanisms that output a
representation of the full posterior distribution.

@@

Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$ given $\dataobs$ is defined as:
\[
  \Pr(\theta | \dataobs) \equiv \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression $\Pr(\dataobs | \theta)$ denotes the
\emph{likelihood} of observing $\dataobs$ under a value of
$\theta$. Since we consider $\dataobs$ to be fixed, the likelihood is
a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$, that is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distribution of the prior. In this case we say that the prior
is conjugate to the likelihood function. Use of a conjugate prior
simplifies calculations and allows for inference to be performed in a
recursive fashion over the data.
% Then, we have:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]

@@

In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
We will consider the situation where the underlying data is drawn from a binomial distribution
($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
the parameter --informally called \emph{bias}-- of a Bernoulli
distributed random variable and $n$ is the number of draws. The
prior distribution over $\theta\in [0,1]$ is going to be a beta
distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
in section \ref{} when the private mechanisms will be described.

@@

The beta-binomial model can be immediately generalized to the Dirichlet-multinomial model, with underlying data multinomially distributed.
The \emph{bias} is represented by the parameter $\vtheta$, the vector of parameters of a categorically distributed random variable. The prior distribution over $\vtheta\in [0,1]^{k}$
is given by a Dirichelet distribution, $\dirichlet{n}{\valpha}$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbb{R}^{+})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where $\iverson{\cdot}$ represents Iverson bracket notation.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots \Delta\alpha_k)$ the posterior distribution over $\vtheta$ turns out to be
\[
   \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)=\dirichlet{\valpha + \Delta \valpha} 
\]
where $+$ denotes the componentwise sum of vectors of reals.

@@

A randomized mechanism $\mathcal{M}$ with domain $\mathbb{N}^{|\mathcal{X}|}$ and range $\range$, is $(\epsilon, \delta)$-differential privacy, if for all adjacent
\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^n$  we say that $\dataobs$ and $\dataobs'$ are adjacent and we write, $\adj{\dataobs}{\dataobs'}$, iff\\
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$. } $\dataobs, \dataobs' \in \mathbb{N}^{|\mathcal{X}|}$ and all $\mathcal{S} \subseteq \range$:
\begin{equation*}
\Pr[\mathcal{M}(\dataobs) \in \mathcal{S}] \leq e^\epsilon \Pr[\mathcal{M}(\dataobs') \in \mathcal{S}].
\end{equation*}

@@

However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as defined above.
\todo{This
accuracy metric is meaningless when the results released are
distributions rather than numerical values. In contrast, distribution
metrics such as $f$-divergence, Hellinger distance, etc. come into
mind overtly when we are measuring distance between
distributions. }
{\color{blue}im not sure this is a good way to put it...}
This gives us motivation on exploring the design space
of mechanisms by considering different norms (a distribution metric)
to compute the sensitivity and provide guarantees on the accuracy.

We will use the Hellinger distance $\hellinger(\cdot,\cdot)$, 
Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the Hellinger distance between them is defined as:
\[
\todo{def}
\]
When $\textup{P}=\betad(\alpha_1, \beta_1)$ and $\textup{Q}=\betad(\alpha_2, \beta_2)$ the above reduces to: 
\[
  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]
